---
title: "Lab 6 Comparing means"
author: "Lukas Robinson"
date: "`r Sys.Date()`"
output: github_document
---

Fill out this report and submit it as your completed assignment.

You can view the R markdown script that created this report here: [README.Rmd](README.Rmd)

## $t$-test

Researchers studying the number of electric fish species living in various parts of the Amazon basin were interested in whether the presence of tributaries affected the local number of electric fish species in the main rivers (Fernandes et al. 2004).

They counted the number of electric fish species above and below the entrance point of a major tributary at 12 different river locations.

The data is provided in your GitHub repository.

For each question below, write a sentence answering the question and show the code you used to come up with the answer, if applicable.

load packages & read data sets

```{r}
library(tidyverse)
fish <- read_csv("chap12q19ElectricFish.csv")
crabs <- read_csv("chap15q27FiddlerCrabFans.csv")
fish_long <- 
  pivot_longer(fish, speciesUpstream:speciesDownstream,
               names_to = "location",
               values_to = "species") %>% 
  mutate(location = str_remove(location, c("species")))
```

### Question A

> Test the hypothesis that the tributaries have no effect on the number of species of electric fish.

```{r t-test}
t.test(formula = species ~ location, data = fish_long)
```

The p-value is greater than 0.5 so we are accepting the null hypothesis.

### Question B

> What is the difference in the mean numbers of species between areas upstream and downstream of a tributary? What is the 95% confidence interval of this difference in means?

The mean difference is 1.83334 and the 95% confidence interval of this difference in means is -4.587031 and 8.253697.

### Question C

> State the assumptions that you had to make to complete parts (A) and (B). Create a graph to assess whether one of those assumptions was met.

The assumptions of a t-test is the sampling units are randomly sampled from populations and the paired differences have a normal distribution in populations. The paired differences can be tested by creating a histogram of the two populations. This histogram then shows that the data does have characteristics of a normal distribution in the sample size fo 12 samples per location.

```{r}
fish_long %>% 
  ggplot(aes(x = species)) +
  geom_histogram(
    aes(fill = location),
    bins = 8,
    alpha = 0.5,
    position = "identity"
  ) +
  scale_fill_manual(values = c("red", "cyan2")) +
  theme_minimal() +
  facet_wrap( ~ location)
```

## ANOVA

Fiddler crabs are so called because males have a greatly enlarged "major" claw, which is used to attract females and to defend a burrow.

Darnell and Munguia (2011) recently suggested that this appendage might also act as a heat sink, keeping males cooler while out of the burrow on hot days.

To test this, they placed four groups of crabs into separate plastic cups and supplied a source of radiant heat (60-watt light bulb) from above. The four groups were intact male crabs, male crabs with the major claw removed; male crabs with the other (minor) claw removed (control); and intact female fiddler crabs.

They measured the body temperature of crabs every 10 minutes for 1.5 hours. These measurements were used to calculate a rate of heat gain for every individual crab in degrees C/log minute. Rates of heat gain for all crabs are provided in the accompanying data file.

### Question D

> Graph the distribution of body temperatures for each crab type:

```{r}
crabs %>% 
  filter(!is.na(crabType)) %>% 
  ggplot(aes(x = bodyTemperature)) +
  geom_histogram(
    aes(fill = crabType),
    bins = 8,
    alpha = 0.5,
    position = "identity"
  ) +
  facet_wrap( ~ crabType, ncol = 1)
```

### Question E

> Does body temperature vary among crab types? State the null and alternative hypothesis, conduct and ANOVA, and interpret the results.
>
> HO: The body temperature is equal among all crab types
>
> HA: At least one crab type is different from the others

```{r}
aov_crab <-
  aov(bodyTemperature ~ crabType, data = crabs)

summary(aov_crab)
```

The ANOVA results that the F value is greater than one and this means we are going to support the alternative hypothesis. The p-value is less than 0.5 which means we are rejecting the null hypothesis meaning there is at least one crab type different from the others.
